The non-random assembly of functional motifs in plant-pollinator networks worldwide

### 2022-01-03 Manuscript/Main_text_Oikos.Rmd
edits for publication

Keyworkds: Plant-pollinator networks, Plant-pollinator interactions, network motifs, functional groups

@@

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled in a larger set of networks. Floral visitor functional groups were divided by the main taxonomic groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) if there is a common invariant structural property in the overall motif networks (i.e. over and under-represented motifs); (ii) which functional groups are over and under-represented in the different motif positions; and, (iii) if there are over and under-reprsented functional group combinations within a motif.

@@

First, plant species were grouped through hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. For this, we used the trait dataset collated in Chapter 1 that comprised a total of 1506 species including the 503 species considered in this study (Supplementary material Chapter 1 **Table S1**). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30$\%$; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species [@dolnicar2014]. To feed the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].

>>>>

Keyworkds: plant-pollinator networks, plant-pollinator interactions, network motifs, functional groups, plant traits, floral visitors

@@

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled in a larger set of networks. Floral visitor functional groups were divided by the main taxonomic groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) if there is a common invariant structural property in the overall motif networks (i.e. over and under-represented motifs); (ii) which functional groups are over and under-represented in the different motif positions; and, (iii) if there are over- and under-represented functional group combinations within a motif.

@@

First, plant species were grouped through hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. For this, we used the trait dataset collated in @lanuza2021 that comprised a total of 1506 species including the 503 species considered in this study (Supplementary material Chapter 1 **Table S1**). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30$\%$; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species [@dolnicar2014]. To feed the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].
